---
title: "Technical Assessment for Data Anaylst (Clinical Quality & Performance Management)"
author: "Kok Jim Meng"
date: "9 November 2021"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
runtime: shiny
---

# 1. Install the Relevant Packages

First of all, I need to install the relevant libraries that are used in this assessment. To install, uncomment the following code chunk.

```{r echo = TRUE, message=FALSE, warning= FALSE }
#install.packages("tidyverse")
#install.packages("glue")
#install.packages("shiny")
#install.packages("ggthemes")
```

# 2. Import

Here, I import the following packages that have installed earlier.

```{r echo = TRUE, message=FALSE, warning= FALSE }
library(tidyverse)
library(glue)
library(shiny)
library(ggthemes)
```

Furthermore, the datasets are imported based on the instructions given and they were downloaded from [https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-11-13](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-11-13)

```{r echo = TRUE, message=FALSE, warning= FALSE }
malaria_deaths_dataset <- read_csv("malaria_deaths.csv")
malaria_deaths_age_dataset <- read_csv("malaria_deaths_age.csv")
malaria_inc <- read_csv("malaria_inc.csv")
```

# 3. Data Processing

```{r echo = TRUE, message=TRUE, warning= FALSE }
#glimpse(malaria_deaths_dataset)
#glimpse(malaria_deaths_age_dataset)
#glimpse(malaria_inc)
#malaria_deaths_dataset$`Deaths - Malaria - Sex: Both - Age: Age-standardized (Rate) (per 100,000 people)`
#is.nan(malaria_deaths_dataset$`Deaths - Malaria - Sex: Both - Age: Age-standardized (Rate) (per 100,000 people)`)
```

`glimpse` function is to allow us to look at the variable names and their types and samples of values.

```{r echo = TRUE, message=TRUE, warning= FALSE }
glimpse(malaria_deaths_dataset)
glimpse(malaria_deaths_age_dataset)
glimpse(malaria_inc)
```

However, both the malaria_deaths_dataset's death variable and malaria_inc's incidence have a long name each. Hence, renaming is conducted to showcase the details of the two datasets.

```{r echo = TRUE, message=TRUE, warning= FALSE }
malaria_deaths_dataset <- malaria_deaths_dataset %>% 
  rename(Deaths = `Deaths - Malaria - Sex: Both - Age: Age-standardized (Rate) (per 100,000 people)`)

malaria_inc <- malaria_inc %>% 
  rename(Incidence = `Incidence of malaria (per 1,000 population at risk) (per 1,000 population at risk)`)

glimpse(malaria_deaths_dataset)
glimpse(malaria_inc)
```

From here, I have selected the relevant variables that will be used to generate the data visualisations.

```{r echo = TRUE, message=TRUE, warning= FALSE }
malaria_deaths_dataset <- malaria_deaths_dataset %>% 
  select(Entity, Year, Deaths)

malaria_deaths_age_dataset <- malaria_deaths_age_dataset %>% 
  select(entity, year, age_group, deaths)

malaria_inc <- malaria_inc %>% 
  select(Entity, Year, Incidence)

glimpse(malaria_deaths_dataset)
glimpse(malaria_deaths_age_dataset)
glimpse(malaria_inc)
```

To ensure meaningful insights in the visualisation section, checking for values with zeros are put into place. This is to ensure that if a country or a region has zero values throughout the years. With that, it does not bring any meaningful insights.

```{r echo = TRUE, message=TRUE, warning= FALSE }
colSums(malaria_deaths_dataset == 0)
```

```{r echo = TRUE, message=TRUE, warning= FALSE }
malaria_deaths_dataset %>% 
  filter(Deaths == 0)
```

```{r echo = TRUE, message=TRUE, warning= FALSE }
colSums(malaria_deaths_age_dataset == 0)
```

```{r echo = TRUE, message=TRUE, warning= FALSE }
malaria_deaths_age_dataset %>% 
  filter(deaths == 0)
```

```{r echo = TRUE, message=TRUE, warning= FALSE }
colSums(malaria_inc == 0)
```

```{r echo = TRUE, message=TRUE, warning= FALSE }
malaria_inc %>% 
  filter(Incidence == 0)
```

Hence, the dependent variable - death - of both the malaria_deaths_dataset and malaria_deaths_age_dataset have zero values through each country's or region's periods. Therefore, a country or region with zero deaths throughout the respective periods does not bring in meaningful insights, the datasets will remove countries or regions with zero deaths throughout the periods.

```{r echo = TRUE, message=FALSE, warning= FALSE }
malaria_deaths_dataset <- malaria_deaths_dataset[ malaria_deaths_dataset$Deaths!=0, ]

malaria_deaths_age_dataset <- malaria_deaths_age_dataset[ malaria_deaths_age_dataset$deaths!=0, ]
```

# 4. Data Visualisation

The following visualisations allow users to select the country or the region, together with the selecting of age groups for 4.2 Malaria Deaths in terms of Country/Region and Age Group visualisation, that he/she is interested in via the dropdown menus.

In this technical assessment, I'll be using Indonesia as an example for the analysis.

## 4.1 Malaria Deaths in terms of Country/Region

```{r echo = FALSE, error = FALSE, message=FALSE}
selectInput("d_countries","Select a Country or a Region",
              choices =  c(unique(as.character(malaria_deaths_dataset$Entity))))

renderPlot({
      malaria_deaths_choosecountry <- malaria_deaths_dataset[ malaria_deaths_dataset$Entity==input$d_countries, ] 
      ggplot(data=malaria_deaths_choosecountry,aes(x=Year,y=Deaths)) + 
        geom_bar(stat="Identity", fill ="red") +
        labs(x = "Year", y = "No. of Deaths (per 100,000 people)", title = paste("Malaria Deaths (per 100,000 people) in", input$d_countries, "from 1990 to 2016", sep=" ")) +
        ggthemes::theme_economist()
    })
```

In the above chart, the progress of malaria deaths (per 100,000) have reduced steadily across the years from 1990 to 2016.

## 4.2 Malaria Deaths in terms of Country/Region and Age Group selected

```{r echo = FALSE, error = FALSE, message=FALSE}
selectInput("da_countries","Select a Country or a Region",
              choices =  c(unique(as.character(malaria_deaths_age_dataset$entity))))
selectInput("agegroup","Select an Age Group",
              choices =  c(unique(as.character(malaria_deaths_age_dataset$age_group))))


renderPlot({
      malaria_deaths_age_choosecountry <- malaria_deaths_age_dataset[ malaria_deaths_age_dataset$entity==input$da_countries, ] 
      malaria_deaths_age_chooseage <- malaria_deaths_age_choosecountry[ malaria_deaths_age_choosecountry$age_group==input$agegroup, ] 
      ggplot(data=malaria_deaths_age_chooseage,aes(x=year,y=deaths)) + 
        geom_bar(stat="Identity", fill = "red") +
        labs(x = "Year", y = "No. of Deaths", title = paste("Malaria Deaths in terms of Aged", input$agegroup, "in", input$da_countries, "from 1990 to 2016", sep=" ")) +
        ggthemes::theme_economist()
    })
```

For aged under 5 years old and aged 5 to 14 years old, similar to the chart in 4.1 Malaria Deaths in terms of Country/Region, the number of malaria deaths has reduced steadily across the years from 1990 to 2016.

However, for aged 15 to 49 years old, the number of malaria deaths rose steadily to the year 1998 and reduce progressively after the year.

Lastly, for both 50 to 69 years old and 70 years old & above, the number of malaria deaths rose steadily to the year 1998 and reduce progressively under 2003 and started to rise again after then.

## 4.3 Malaria Incidence

```{r echo = FALSE, error = FALSE, message=FALSE}
selectInput("i_countries","Select a Country or a Region",
              choices =  c(unique(as.character(malaria_inc$Entity))))

renderPlot({
      malaria_inc_choosecountry <- malaria_inc[ malaria_inc$Entity==input$i_countries, ] 
      ggplot(data=malaria_inc_choosecountry,aes(x = Year, y = Incidence)) + 
  geom_line(color="red") +
        geom_point(color="red") +
        labs(x = "Year", y = "Malaria Incidence", title = paste("Malaria Incidence (per 1,000 population at risk) in", input$i_countries, sep=" ")) +
        ggthemes::theme_economist()
    })
```

In terms of malaria incidence (per 1,000 population at risk), it has increased at a slower rate from 2000 to 2010 in Indonesia. After 2010, it has dropped sharply in 2015.